Approximation as Differentiation

### 2025-05-05 appendix/notes/auto-diff.tex
smooth -> differentiable (don't need C^\infinity functions).

\begin{definition}[Forward mode automatic differentiation functor]
Define the functor $\tangents_*$ which sends every vector space $\RR^n$ to itself and every smooth map $f:
\RR^m \to \RR^n$ to the function $\tangents_*(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times (\RR^m
\linearto \RR^n)$, associating to every point its image in $f$ and the forward derivative of $f$ at that
point.
\end{definition}

\begin{definition}[$\Diff$]
\label{def:auto-diff:Diff}
Write $\Diff_*$ for the category where the objects are all finite vector spaces $\RR^n$ and the morphisms are
all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Writing $f_1$ for $\pi_1 \comp f$ and $f_2$ for
$\pi_2 \comp f$, composition in $\Diff_*$ is given by:
\begin{align*}
(g \comp f)_1(x) &= f_1(g_1(x)) \\
(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
\end{align*}
\end{definition}

@@

\begin{definition}[Reverse mode automatic differentiation functor]
Define the functor $\tangents^*$ which also sends every vector space $\RR^n$ to itself but which sends every
smooth map $f: \RR^m \to \RR^n$ to the function $\tangents^*(f) = x \mapsto (f(x), \pullf{f}_x): \RR^m \to
\RR^n \times (\RR^n \to \RR^m)$.
\end{definition}

@@

If we view any $\Diff$ object $\RR^n$ (\defref{auto-diff:Diff}) as implicitly paired with a constant family
$\{\RR^n\}_{x \in \RR^n}$ and any $\Diff$ morphism $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the
pair of a smooth function $f_1: \RR^m \to \RR^n$ and constant family of linear maps $f_2: \{\delta f_x: \RR^m
\linearto \RR^n\}_{x \in \RR^m}$, then it is clear that $\Diff$ is equivalent to a full subcategory of
$\Fam(\FinVect)$.

>>>>

\begin{definition}[Forward mode automatic differentiation functor]
Define the functor $\tangents_*$ which sends every vector space $\RR^n$ to itself and every differentiable map
$f: \RR^m \to \RR^n$ to the function $\tangents_*(f) = x \mapsto (f(x), \pushf{f}_x): \RR^m \to \RR^n \times
(\RR^m \linearto \RR^n)$, associating to every point its image in $f$ and the forward derivative of $f$ at
that point.
\end{definition}

For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ write $f_1$ for $\pi_1 \comp f$ and $f_2$ for
$\pi_2 \comp f$.

\begin{definition}[$\Diff$]
\label{def:auto-diff:Diff}
Define $\Diff_*$ to be the category where the objects are all finite vector spaces $\RR^n$ and the morphisms
are all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, with $f_1$ differentiable. Composition in
$\Diff_*$ is given by:
\begin{align*}
(g \comp f)_1(x) &= f_1(g_1(x)) \\
(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
\end{align*}
\end{definition}

@@

\begin{definition}[Reverse mode automatic differentiation functor]
Define the functor $\tangents^*$ which also sends every vector space $\RR^n$ to itself but which sends every
differentiable map $f: \RR^m \to \RR^n$ to the function $\tangents^*(f) = x \mapsto (f(x), \pullf{f}_x): \RR^m
\to \RR^n \times (\RR^n \to \RR^m)$.
\end{definition}

@@

Consider the functor $\Diff \hookrightarrow \Fam(\FinVect)$ which pairs objects $\RR^n$ with the constant
family $\{\RR^n\}_{x \in \RR^n}$ and sends morphisms $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ to
the pair of a differentiable function $f_1: \RR^m \to \RR^n$ and constant family of linear maps $f_2: \{\delta
f_x: \RR^m \linearto \RR^n\}_{x \in \RR^m}$. Then it is clear that $\Diff$ is equivalent to a full subcategory
of $\Fam(\FinVect)$. For any morphisms $(f, \partial f): (\RR^m, \{\RR^m\}_{x \in \RR^m}) \to (\RR^n,
\{\RR^n\}_{y \in \RR^n})$ and $(g, \partial g): (\RR^n, \{\RR^n\}_{y \in \RR^n}) \to (\RR^k, \{\RR^k\}_{z \in
\RR^k})$ in the image of that inclusion functor, we have for any $x \in \RR^m$:
\begin{itemize}
\item $\partial f_x: \RR^m \to \RR^n$;
\item $\reindex{\partial g}{f}_x = \partial g_{f(x)}: \RR^n \to \RR^k$
\end{itemize}
